---
title: "Live Session Assignment 9"
author: "Kevin Cannon"
date: "November 13, 2016"
output: 
  html_document:
       keep_md: yes
---

# Dataset
```{r}
# read in file
file <- "https://raw.githubusercontent.com/merdirafiei/CHF/master/CHF.csv"
chf <- read.csv(file)
# explore data
head(chf)
str(chf)
#attach dataframe
attach(chf)
```

### 1. Show the mean for Amount Paid by Medicare, Total Accommodation Charges, and Total Departmental Charges
```{r}
mean(AmtReim)
mean(TotAccomChg)
mean(TotDeptChg)
```

### 2. Show the standard deviation for Amount Paid by Medicare, Total Accommodation Charges, and Total Departmental Charges
```{r}
sd(AmtReim)
sd(TotAccomChg)
sd(TotDeptChg)
```


### 3. Use tapply() to compute the mean for Amount Paid by Medicare, Total Accommodation Charges, and Total Departmental Charges, by gender
```{r}
# sex = 1 for male
# sex = 2 for female
tapply(AmtReim, Sex, mean)
tapply(TotAccomChg, Sex, mean)
tapply(TotDeptChg, Sex, mean)
```

### 4. Create a boxplot for Amount Paid by Medicare, Total Accommodation Charges, and Total Departmental Charges by gender
```{r}
# medicare reimbursement
boxplot(AmtReim~Sex, main=toupper("Medicare Reimbursement by Gender"))

# total accomodation charges
boxplot(TotAccomChg~Sex, main=toupper("Total Accomodation Charges by Gender"))

# total departmental charges
boxplot(TotDeptChg~Sex, main=toupper("Total Departmental Charges by Gender"))
```

### 5. Create a histogram for the following variables and describe what you see for each: 
#### a. Accommodation charges
#### b. Gender
#### c. Age category

```{r}
# accomodation charges
hist(TotAccomChg, main=toupper("Total Accomodation Charges"))
# gender
hist(Sex, main=toupper("Gender"))
# category
hist(Age, main=toupper("Ages"))
```

### 6. Graph the count of each gender for each admission source using a bar plot
```{r}
admsrcBYsex <- table(Sex, admsrc)

#create barplot
barplot(admsrcBYsex, main = "Admissions Source by Gender", xlab = "Admissions Source Code", legend = c("Male","Female"), beside=TRUE)
```

### 7. Using a bar plot, plot the mortality rate for MS-DRG 292 and MS-DRG 293
```{r}
newDRG292 <- subset(chf, drgcode==292)

# discharge destination is expired
newDRG292$dischdest[newDRG292$dischdest==20]

# expired/all discharge destinations
mortrate292 <-round(length(newDRG292$dischdest[newDRG292$dischdest==20])/length(newDRG292$dischdest),3)

# subset of DRG 292 codes 
newDRG293 <- subset(chf, drgcode==293)

# discharge destination is expired
newDRG293$dischdest[newDRG293$dischdest==20]

# expired/all discharge destinations
mortrate293 <-round(length(newDRG293$dischdest[newDRG293$dischdest==20])/length(newDRG293$dischdest),3)

#create table 
mortnames <- c("Mortality DRG 292", "Mortality DRG 293")
mortrate <- c(mortrate292, mortrate293)
morttable <- table(mortnames,mortrate)

# bar graph of mortality rates of DRG = 292 and DRG = 293
barplot(mortrate, main = "Mortality Rate among DRG codes 292 and 293", xlab = "DRG Code", legend = c("292", "293"), beside=TRUE)
```

### 8. You must determine if the average length of stay (LOS) for patients with MS-DRG
```{r}
# 1-sample ttest
# LOS for drg 291
LOS291 <- LOS[drgcode == 291]

# ttest on mu =6
t.test(LOS291, mu=6)
```

From the one-sample t-test, there is enough evidence to reject the null hypothesis that the mean length of stay for DRG 291 is six days. The results of the t-test suggest that the mean length of stay for DRG 291 is in fact greater than the average stay length.